---
title: "Earnings Calls and Stock Analysis"
author: "Raunak Sood"
date: "2023-06-08"
output: html_document
---

```{r setup, include=FALSE}
# Load required packages
library(tidytext)
library(dplyr)
library(ggplot2)
library(lubridate)
library(purrr)
library(stargazer)
library(tidyverse)



# Folder path for the text files
folder_path <- "/Users/raunaksood/Downloads/dataverse_files/Transcripts"

# List of company abbreviations
companies <- c("MU", "CSCO", "NVDA", "AMZN", "AAPL", "MSFT", "GOOGL", "AMD", "ASML", "INTC")

# Create a list to store the transcript data
transcript_data <- list()

# Read the text files for each company and store in the list
for (company in companies) {
  # List of text file names for the current company
  file_names <- list.files(path = file.path(folder_path, company), pattern = "\\.txt$", full.names = TRUE)
  
  # Create a list to store the transcript data for the current company
  company_transcripts <- list()
  
  
  
  # Read each file and store its contents in the list
  for (file_name in file_names) {
    transcript <- readLines(file_name)
    company_transcripts[[basename(file_name)]] <- unlist(transcript)
  }
  
  # Store the transcript data for the current company in the main list
  transcript_data[[company]] <- company_transcripts
}

# Accessing the transcript data for a specific company (e.g., "MU")
mu_transcript <- transcript_data[["MU"]]
cisco_transcript <- transcript_data[["CSCO"]]
nvda_transcript <- transcript_data[["NVDA"]]
amzn_transcript <- transcript_data[["AMZN"]]
aapl_transcript <- transcript_data[["AAPL"]]
msft_transcript <- transcript_data[["MSFT"]]
googl_transcript <- transcript_data[["GOOGL"]]
amd_transcript <- transcript_data[["AMD"]]
asml_transcript <- transcript_data[["ASML"]]
intc_transcript <- transcript_data[["INTC"]]

file_path <- "/Users/raunaksood/Downloads/dataverse_files/Stock Values and Sector Index"


# Read the stock files for the companies
nvda_data <- read.csv(file.path(file_path, "NVDA.csv"))
nvda_data$Company <- "NVDA"

mu_data <- read.csv(file.path(file_path, "MU.csv"))
mu_data$Company <- "MU"

msft_data <- read.csv(file.path(file_path, "MSFT.csv"))
msft_data$Company <- "MSFT"

intc_data <- read.csv(file.path(file_path, "INTC.csv"))
intc_data$Company <- "INTC"

googl_data <- read.csv(file.path(file_path, "GOOGL.csv"))
googl_data$Company <- "GOOGL"

csco_data <- read.csv(file.path(file_path, "CSCO.csv"))
csco_data$Company <- "CSCO"

asml_data <- read.csv(file.path(file_path, "ASML.csv"))
asml_data$Company <- "ASML"

amzn_data <- read.csv(file.path(file_path, "AMZN.csv"))
amzn_data$Company <- "AMZN"

amd_data <- read.csv(file.path(file_path, "AMD.csv"))
amd_data$Company <- "AMD"

aapl_data <- read.csv(file.path(file_path, "AAPL.csv"))
aapl_data$Company <- "AAPL"


# Create the all_stock_data dataframe
all_stock_data <- data.frame(Date = character(), stringsAsFactors = FALSE)

# Merge all the dataframes into all_stock_data
all_stock_data <- rbind(nvda_data, mu_data, msft_data, intc_data, googl_data, csco_data, asml_data, amzn_data, amd_data, aapl_data)

# Order the dataframe by company and date
all_stock_data <- all_stock_data[order(all_stock_data$Company, all_stock_data$Date), ]

# Function to convert transcript data to data frame
convert_to_dataframe <- function(transcript) {
  transcript_df <- data.frame(text = unlist(transcript), stringsAsFactors = FALSE)
  return(transcript_df)
}

# Convert transcript data to data frames
MU_transcript_df <- convert_to_dataframe(mu_transcript)
CSCO_transcript_df <- convert_to_dataframe(cisco_transcript)
NVDA_transcript_df <- convert_to_dataframe(nvda_transcript)
AMZN_transcript_df <- convert_to_dataframe(amzn_transcript)
AAPL_transcript_df <- convert_to_dataframe(aapl_transcript)
MSFT_transcript_df <- convert_to_dataframe(msft_transcript)
GOOGL_transcript_df <- convert_to_dataframe(googl_transcript)
AMD_transcript_df <- convert_to_dataframe(amd_transcript)
ASML_transcript_df <- convert_to_dataframe(asml_transcript)
INTC_transcript_df <- convert_to_dataframe(intc_transcript)

# Create a list of transcript data frames
transcript_dfs <- list(
  MU = convert_to_dataframe(mu_transcript),
  CSCO = convert_to_dataframe(cisco_transcript),
  NVDA = convert_to_dataframe(nvda_transcript),
  AMZN = convert_to_dataframe(amzn_transcript),
  AAPL = convert_to_dataframe(aapl_transcript),
  MSFT = convert_to_dataframe(msft_transcript),
  GOOGL = convert_to_dataframe(googl_transcript),
  AMD = convert_to_dataframe(amd_transcript),
  ASML = convert_to_dataframe(asml_transcript),
  INTC = convert_to_dataframe(intc_transcript)
)

# Combine all transcript data frames into a single data frame
all_transcripts_df <- bind_rows(transcript_dfs, .id = "company")

# Remove ".txt" and random numbers from row indices and store in a new "Date" column
all_transcripts_df$Date <- gsub("\\.txt.*", "", row.names(all_transcripts_df))

all_transcripts_df$Date <- gsub("-[^-]*$", "", all_transcripts_df$Date)



# Perform sentiment analysis using the "bing" lexicon
sentiment_analysis <- all_transcripts_df %>%
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(company, sentiment) %>%
  mutate(sentiment = ifelse(sentiment == "positive", "Positive", "Negative"))


# Top 10 most common words
top_words <- all_transcripts_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) %>%
  top_n(10)


# Convert the "Date" column to the appropriate format
all_stock_data$Date <- as.Date(all_stock_data$Date)

# Convert the "Date" column in all_transcripts_df to the appropriate format
all_transcripts_df$Date <- lubridate::ymd(all_transcripts_df$Date)

# Extract the quarter information from the "Date" column
all_stock_data$Quarter <- paste0("Q", quarter(all_stock_data$Date), " ", year(all_stock_data$Date))

# Extract the quarter information from the "Date" column in all_transcripts_df
all_transcripts_df$Quarter <- paste0("Q", quarter(all_transcripts_df$Date), " ", year(all_transcripts_df$Date))


# Calculate sentiment scores for each transcript
sentiment_scores <- all_transcripts_df %>%
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  mutate(sentiment = ifelse(sentiment == "positive", "Positive", "Negative")) %>%
  count(company, Quarter, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment_index = Positive - Negative)

# Merge sentiment scores with stock data
merged_data <- merge(all_stock_data, sentiment_scores, by.x = c("Company", "Quarter"), by.y = c("company", "Quarter"), all.x = TRUE)



# Select the relevant columns from merged_data for positive_data
positive_data <- merged_data %>%
  select(Company, Quarter, Close, Positive)

# Convert the "Date" column to the appropriate format
all_stock_data$Date <- as.Date(all_stock_data$Date)
all_transcripts_df$Date <- as.Date(all_transcripts_df$Date)

merged_data <- merge(all_stock_data, sentiment_scores, by.x = c("Company", "Quarter"), by.y = c("company", "Quarter"), all.x = TRUE)

relevant_data <- merged_data %>% select(Company, Quarter, Close, sentiment_index)

relevant_data <- relevant_data %>%
  group_by(Company) %>%
  mutate(stock_price_change = lead(Close) - lag(Close))

correlation_results <- relevant_data %>%
  group_by(Company) %>%
  summarize(correlation = cor(sentiment_index, stock_price_change, use = "pairwise.complete.obs"))

# Calculate sentiment scores for each transcript by quarter
sentiment_scores_quarter <- all_transcripts_df %>%
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  mutate(sentiment = ifelse(sentiment == "positive", "Positive", "Negative")) %>%
  count(company, Quarter, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment_index = Positive - Negative) %>%
  mutate(Quarter = paste0("Q", Quarter, " "))

```

This report aims to uncover the effects of earnings calls on stock prices. In this report we look at data of 10 of the largest companies in 2020: Micron Technology Inc. (MU), Cisco Systems Inc. (CSCO), NVIDIA Corporation (NVDA), Amazon.com Inc. (AMZN), Apple Inc. (AAPL), Microsoft Corporation (MSFT), Alphabet Inc. or Google (GOOGL), Advanced Micro Devices Inc. (AMD), ASML Holding NV (ASML), Intel Corporation (INTC). Earnings call transcripts and stock data will be analyzed from 2016-2020. This analysis looks at stock trends, text and sentiment analysis, and effects of earnings calls sentiment on stock price.  


```{r plot1, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Calculate the initial closing price for each stock
initial_prices <- all_stock_data %>%
  filter(year(Date) == 2016) %>%
  group_by(Company) %>%
  summarize(initial_price = first(Close))

# Merge the initial prices with the all_stock_data
merged_data <- merge(all_stock_data, initial_prices, by = "Company")

# Normalize the stock prices to 100
merged_data$Close_normalized <- (merged_data$Close / merged_data$initial_price) * 100

# Line plot of normalized stock prices over time
normalized_plot <- ggplot(merged_data, aes(x = Date, y = Close_normalized, color = Company)) +
  geom_line() +
  labs(x = "Time", y = "Normalized Stock Price", title = "Figure 1: Normalized Stock Prices Over Time") +
  theme_minimal()

# Display the plot
print(normalized_plot)

```

In Figure 1, we see a normalized stock plot of all ten companies between 2016 and 2020. Each stock price is normalized at a 100 dollars so we get a better look at the growth of different companies. We find that AMD and NVDA exhibit the largest growth, although all companies show some growth. 


```{r plot3, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Plotting sentiment scores by company
sentiment_plot <- ggplot(sentiment_scores, aes(x = reorder(company, -sentiment_index), y = sentiment_index, fill = company)) +
  geom_col() +
  labs(x = "Company", y = "Sentiment Index", title = "Figure 2: Sentiment Index by Company") +
  scale_fill_manual(values = c("#F8766D", "#00BFC4", "#FF61A6", "#7CAE00", "#B2D732", "#ECD078", "#E58606", "#2B6BAC", "#CAB3AB", "#6495ED")) +
  theme_minimal() +
  theme(legend.position = "none")

# Display the plot
print(sentiment_plot)

```

In Figure 2, we look at sentiment index of each company. This index is calculated by positive words minus negative words. We see that earnings calls primarily consist of positive words, with Google having the highest count. 


```{r plot4, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Top 10 most common words (excluding insignificant words and numbers)
top_words <- all_transcripts_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  filter(!word %in% c("quarter", "call", "question", "analyst", "time", "operator", "ceo", "id", "division")) %>%
  filter(!str_detect(word, "^\\d+$")) %>%
  count(word, sort = TRUE) %>%
  top_n(10)

# Bar plot of the modified top words
word_plot <- ggplot(top_words, aes(x = reorder(word, n), y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Word", y = "Frequency", title = "Figure 3: Top 10 Most Common Words (Excluding Insignificant Words)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the modified plot
print(word_plot)
```

Figure 3 looks at most common words found in earnings calls. Excluded words include words like ceo, quarter, and call, which are insignificant to this analysis. Words such as revenue, growth, and costumers, all make sense to indicate a growing business to analysts and the public. 


```{r plot11, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Define non-relevant words to exclude
non_relevant_words <- c("quarter", "analyst", "ceo", "gross", "id")


# Get the top 300 most common words, excluding non-relevant words and numbers
top_words <- all_transcripts_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  filter(!(word %in% non_relevant_words)) %>%
  filter(!grepl("\\d", word)) %>%
  count(word, sort = TRUE) %>%
  top_n(1000)

# Perform sentiment analysis on the top 300 words using the "afinn" lexicon
sentiment_scores <- top_words %>%
  inner_join(get_sentiments("afinn"), by = c("word" = "word")) %>%
  filter(!is.na(value))

# Select the top 20 words with sentiment values assigned, 5 positive and 5 negative
positive_words <- sentiment_scores %>%
  filter(value > 0) %>%
  top_n(5, wt = n) %>%
  mutate(sentiment_label = "Positive")

negative_words <- sentiment_scores %>%
  filter(value < 0) %>%
  top_n(5, wt = n) %>%
  mutate(sentiment_label = "Negative")

top_words <- bind_rows(positive_words, negative_words)

# Create the histogram
word_histogram <- ggplot(top_words, aes(x = reorder(word, sentiment_label), y = n, fill = sentiment_label)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Positive" = "blue", "Negative" = "red"), guide = FALSE) +
  labs(x = "Word", y = "Count", title = "Figure 4: Top 10 Positive/Negative Words") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))  # Adjust spacing

# Display the histogram
print(word_histogram)

```

In Figure 4, a bar plot displays the five most common positive words, and 5 most common negative words. It is important to note that positive words are much more common than negative words, so these are not the ten most common words. One interesting point that this plot displays is that the most common negative words do not spark extremely negative emotions. 


```{r plot8, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}

# Add a title for the regression analysis
# Run regression analysis
regression_model <- lm(Close ~ Positive, data = positive_data)

# Print regression results with stargazer
stargazer(regression_model, title = "Table 1: Regression Analysis", type = "text")
```
In this regression analysis, we have the close stock price as the outcome variable and the positive sentiment as the explanatory variable. To further explain this regression, it looks at earnings call "x" in year "x" of company"x". It then looks at the change in the stock price at the end of the current quarter from the end of the last quarter. This gives an accurate representation of how earnings calls sentiment may affect stock price. Here we see that they are significantly inversely correlated. This is an odd result, likely due to so many other factors more strongly affecting stock price. Also perhaps the stock price change is taken over too long of an interval for earnings calls sentiment to greatly affect it.    


```{r plot6, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Correlation between sentiment index and stock prices
correlation_plot <- ggplot(correlation_results, aes(x = Company, y = correlation)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(x = "Company", y = "Correlation", title = "Figure 5: Correlation between Sentiment Index and Stock Prices") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))

# Display the plot
print(correlation_plot)
```

Figure 5 is a plot showing the correlation between sentiment and stock price. Once is again we see a mixed correlation likely due to reasons stated above. Microsoft has the highest correlation while Amazon has the least.     

```{r plot9, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Time series of sentiment index
sentiment_time_series <- relevant_data %>%
  filter(Company == "AAPL") %>%
  ggplot(aes(x = Quarter, y = sentiment_index, group = 1)) +
  geom_line(color = "steelblue") +
  geom_point(color = "steelblue", size = 2) +
  labs(x = "Quarter", y = "Sentiment Index", title = "Figure 6: Sentiment Index Over Time (AAPL)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the time series plot
print(sentiment_time_series)
```

Figure 6 looks at the sentiment index of Apple every quarter from 2016-2020. This was conducted as perhaps companies and executives are beginning to pay more close attention to using positive words in calls. This visual appears to back this hypothesis up. 

```{r plot10, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}
# Filter the stock data for Apple (AAPL) company
aapl_stock_data <- all_stock_data[all_stock_data$Company == "AAPL", ]

# Line plot of Apple stock prices over time
aapl_stock_plot <- ggplot(aapl_stock_data, aes(x = Date, y = Close)) +
  geom_line() +
  labs(x = "Date", y = "Stock Price", title = "Figure 7: Apple (AAPL) Stock Prices Over Time") +
  theme_minimal()

# Display the plot
print(aapl_stock_plot)

```

Figure 7 gives us another look at the stock price of Apple over time. This plot appears to somewhat mimic the increase in sentiment index of Apple's earnings calls. This would hint to a correlation between sentiment and true growth in stock price, however the other visualizations and data in this report show otherwise. 


```{r plot12, echo=FALSE, comment=NA, message=FALSE, warning=FALSE}

# Manually set x-axis labels
x_labels <- c("Q1 2016", "Q2 2016", "Q3 2016", "Q4 2016", "Q1 2017", "Q2 2017", "Q3 2017", "Q4 2017",
              "Q1 2018", "Q2 2018", "Q3 2018", "Q4 2018", "Q1 2019", "Q2 2019", "Q3 2019", "Q4 2019",
              "Q1 2020", "Q2 2020", "Q3 2020", "Q4 2020")

# Create a scatter plot with loess line
sentiment_scatter_plot <- ggplot(sentiment_scores_quarter, aes(x = Quarter, y = sentiment_index, color = company)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "black") +
  labs(x = "Quarter", y = "Sentiment Index", title = "Figure 8: Sentiment Index of Companies Over Time") +
  theme_minimal() +
  theme(legend.position = "right", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(labels = x_labels)

# Display the plot
print(sentiment_scatter_plot)

```

Finally, figure 8 is a sentiment index of all companies over time. Here we see that there is no real increase in sentiment for companies over time. This primarily dis-proves the hypothesis made with Figure 6 that executives, as a whole, are changing there verbage to add more positivity. 

To conclude, this report provides insights regarding common words found in earnings calls, different positive and negative words, sentiment indexes, and preliminary analysis regarding effects of earnings calls sentiment on stock price. To conduct this analysis, transcripts and stock data is looked at for 2016-2020 of 10 of the most prominent companies. Originally, the hypothesis made was that a more positive sentiment would be positively correlated with stock price. The analysis conducted in this report shows otherwise. However, it is important to note that other factors such as supply and demand and consumer perception may be larger drivers affecting stock price. Thus, there is omitted variable bias. When re-visiting this analysis, looking at immediate change in stock price after the earnings call takes place will be insightful. 



